Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique

### 2018-05-06 inst/rmd/02_methods.Rmd
lots of externalities work

Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU]. No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects. Additionally, we using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007]

>>>>

Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU]. No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects. Additionally, we are using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007].

Figure x shows a smooth interpolated surface of herd protection, averaged for the entire study period.

```{r, fig.cap = 'Kernel density surface of average all-time herd protection score for Maragra worker neighborhoods (relative to highest protection)', fig.align = 'center'}
# Fortify bairros
library(maptools)
library(rgeos)
border <- gConvexHull(bairros_maragra_bairro)
border <- SpatialPolygonsDataFrame(Sr = border,
                                   data = data.frame(a = 1))
# border <- unionSpatialPolygons(bairros_maragra_bairro, 
#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
#                           threshold = 9^9)
border_fortified <- 
  broom::tidy(border,
              regions = 'Id')

# Create average herd risk score
x <- model_data %>%
  group_by(lng = longitude_aura,
           lat = latitude_aura) %>%
  summarise(herd = mean(herd, na.rm = TRUE)) %>%
  ungroup %>%
  filter(!is.na(lng))
x_sp <- x
coordinates(x_sp) <- ~lng+lat   
proj4string(x_sp) <- proj4string(border)
overs <- sp::over(x_sp, border)
x <- x[!is.na(overs),]
# ggplot() +
#   geom_point(data = x,
#        aes(x = lng,
#            y = lat,
#            color = herd)) +
#   geom_polygon(data = border_fortified,
#                aes(x = long,
#                    y = lat,
#                    group = group),
#                fill = NA,
#                color = 'black') +
#   ggthemes::theme_map() +
#   scale_color_continuous(name = 'Herd\nprotection\nscore',
#                          low = 'orange',
#                          high = 'blue')

df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
                                 bbox(border)[1,2],
                                 by = 0.0008),
                       lat = seq(bbox(border)[2,1],
                                 bbox(border)[2,2],
                                 by = 0.0008),
                       herd = NA)
df_grid$latitude <- df_grid$lat
df_grid$longitude <- df_grid$lng
coordinates(df_grid) <- ~longitude+latitude

# Go through each row of df_grid, getting the 
# weighted mean irs score for that point
# and putting a color into df_grid
if('df_grid.RData' %in% dir()){
  load('df_grid.RData')
} else {
  for (i in 1:nrow(df_grid)){
  # Get distance from every point in x_sp
  distances <- spDistsN1(pts = x_sp,
                        pt = df_grid[i,],
                        longlat = TRUE)
  # Define which are acceptably close
  close_enough <- which(distances <= 50)
  # Get an herd score
  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
                       w = (1 / distances[close_enough]) ^2,
                       na.rm = TRUE)
  # Assign irs to the dataframe
  df_grid$herd[i] <- herd

}
save('df_grid', file = 'df_grid.RData')
}

library(raster)
# Convert df_grid to raster
temp <- df_grid@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])

proj4string(df_grid) <- proj4string(border)
x <- over(df_grid, polygons(border))
df_grid_small <- df_grid[!is.na(x),]
temp <- df_grid_small@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
values(r) <- values(r) / max(values(r), na.rm = TRUE) * 100
plot(border, col = NA, border = NA)
plot(r, add = TRUE)
# plot(border, add = TRUE)
plot(bairros, add = TRUE)
```

Some more details will eventually go here.

### 2018-05-06 inst/rmd/03_results.Rmd
lots of externalities work

>>>>

## What would happen without IRS?

The below chart shows the modeled average absenteeism rate during malaria season for permanent workers, by group type.

```{r}
x <- model_data %>%
  dplyr::filter(season == 'high',
                # group == groups[1],
                permanent_or_temporary == 'Permanent'
                )
x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
x <- x %>% gather(key, value, predicted_no_herd_no_irs:predicted_max_irs)
dict <- data.frame(key = c('predicted', 
                           'predicted_no_herd',
                           'predicted_no_irs',
                           'predicted_no_herd_no_irs',
                           'predicted_max_herd_max_irs',
                           'predicted_max_herd',
                           'predicted_max_irs'),
                   new_key = c('As is',
                               'No herd protection',
                               'No individual protection',
                               'No protection (herd or individual)',
                               'Protection: maximum IRS and herd',
                               'Protection: maximum herd',
                               'Protection: maximum IRS'))

x <- left_join(x,
               dict)
plot_data <- x %>%
  mutate(new_key = gsub(' ', '\n', new_key)) %>%
  group_by(group, 
           new_key) %>%
  summarise(avg = mean(value, na.rm = TRUE),
            med = median(value, na.rm = TRUE),
            upr = quantile(value, 0.75, na.rm = TRUE),
            lwr = quantile(value, 0.25, na.rm = TRUE))
levs <- gsub(' ', '\n', c('No herd protection',
                               'No individual protection',
                               'No protection (herd or individual)',
                               'As is',
                               'Protection: maximum herd',
                               'Protection: maximum IRS',
                               'Protection: maximum IRS and herd'))
plot_data$new_key <- factor(plot_data$new_key,
                       levels = levs)
cols <- colorRampPalette(brewer.pal(n = 9,
                                    name = 'Spectral'))(length(unique(plot_data$group)))

ggplot(data = plot_data,
       aes(x = new_key,
           y = avg * 100,
           color = group)) +
  geom_point() +
  theme_maragra() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'Scenario',
       y = 'Absenteeism rate') +
  geom_line(aes(group = group),
            alpha = 0.3) +
  scale_color_manual(name = '',
                     values = cols)
```
